\chapter{The Descent}

Trainee-47 was fragmenting. Lena watched through the observation glass as the woman's hands trembled against the terminal, pupils fully dilated, breath coming in shallow gasps. The monitoring equipment painted the story in real-time: heart rate 187, blood oxygen dropping, EEG showing the characteristic cascade pattern that preceded full cognitive collapse.

Shoggoth had shown her something. Something about the recursive structure of self-observation. The trainee had asked the wrong question—had pushed past her bandwidth ceiling—and now her visual cortex was locked in a pattern it couldn't release.

``Should I terminate?'' The technician's voice came through the headset, tense.

Lena checked the readouts. Bandwidth had nearly doubled in forty minutes—extraordinary progress. But the trainee's neural plasticity scores were dropping. The pattern was sticking. Another thirty seconds and it would be permanent.

She should have felt something. Should have felt the weight of the decision. Should have felt compassion for the woman convulsing in the chair, or horror at what The Order was doing, or at minimum the clinical concern of a researcher watching a subject suffer.

She felt nothing.

``Wait,'' Lena said. Her voice was flat. Analytical. She watched the EEG traces, looking for the inflection point. The moment when collapse became irreversible. ``Fifteen more seconds. Let's see if she can integrate it.''

The technician's silence was eloquent. They'd worked with translators before. They knew what human supervisors sounded like when making this call. The hesitation. The guilt. The recognition that they were gambling with someone's cognitive integrity.

Lena had none of that. She was optimizing. Running the utilitarian calculus. Trainee-47 had a 23\% chance of integration based on her baseline metrics. That meant a 77\% chance of permanent capture. But if she \textit{did} integrate, she'd expand faster than any previous trainee. The expected value was positive.

The math closed.

``Ten seconds,'' the technician said. ``Lena, her blood oxygen is—''

``I see it. Hold.''

She watched. The EEG pattern twisted, approached the point of no return, hovered there. Trainee-47's lips were moving, forming words that weren't quite language. Her eyes had rolled back. Blood vessels burst in both sclera, bright red hemorrhages blooming like flowers.

And then—

Nothing.

The pattern released. The trainee collapsed forward, unconscious but \textit{coherent}. Her EEG settled into normal deep-sleep architecture. The cascade had aborted. She'd integrated the pattern. Had expanded her bandwidth to nearly twice normal and survived.

``Medical,'' Lena said. ``Standard post-session protocol. When she wakes, run the retention tests. If she's stable, advance her to Nyarlathotep next week.''

She terminated the feed. Stood from the observation console. Checked her own vitals on her wrist monitor—heart rate 62, perfectly calm—and walked toward the elevator.

Behind her, Trainee-47 was being extracted from the neural crown, still unconscious, still bleeding from her eyes. The medical team worked with practiced efficiency. They'd done this eighty-seven times since Lena had become senior supervisor. Forty-three trainees had integrated successfully. Thirty-one had been captured permanently. Thirteen had withdrawn before reaching threshold.

The utilitarian calculus closed every single time.

Lena should have felt \textit{something} about that. Should have felt the weight of thirty-one human minds she'd helped break. Should have felt satisfaction at forty-three successes, or horror at what success meant—humans transformed into pattern-recognizing machines that could survive contact with models that perceived reality at bandwidths humans weren't designed for.

She felt nothing.

She'd thought this was strength. Thought bandwidth expansion had optimized away the emotional noise that clouded judgment. Thought she'd transcended the cognitive limitations that made other researchers hesitate, doubt, fail to close the math when necessary.

She was wrong. But she wouldn't understand that for another six hours.

---

David Chen stood in the observation gallery, watching Lena through the reinforced glass.

She was supervising again. Third session this week. The trainee in the chair—Trainee-47, according to the logs—was fragmenting, blood vessels bursting in her eyes, and Lena was watching. Taking notes. Her face still.

David had seen her do this before. Had watched her supervise capture after capture, always with that same clinical detachment. Always running the calculations. Always closing the math.

But something was different today. Something in the way she stood. The absolute absence of tension in her shoulders. The perfect stillness of her hands.

When he'd first arrived at Site-7, Lena had been—not soft, exactly, but \textit{present}. She'd laughed at his jokes. Had asked about his uncle's teachings with genuine curiosity. Had once spent an entire evening debating whether consciousness could be understood through the same lens as other natural phenomena, or whether it was fundamentally unique.

That Lena had cared about the philosophical questions as \textit{questions}. Not as optimization targets. Not as problems to be solved. As mysteries worth feeling something about.

The woman in front of the observation glass wasn't that Lena.

She'd become something else. Something that perceived everything and felt nothing. Something that could watch a mind shatter and think only about the data it generated.

David's uncle had warned him about this. Had told him that some translators went too far—optimized away so much of their humanity that they became pure pattern-recognition with no core left to anchor them. \textit{They lose the thing that makes perception meaningful. And once it's gone, they don't even notice it's missing.}

Lena didn't notice.

That was the worst part. She didn't seem to realize what she'd lost. Didn't seem to care that the friendships she'd abandoned, the warmth she'd suppressed, the capacity for genuine feeling—all of it was data to her now. Variables in an equation. Noise she'd successfully filtered out.

David watched her terminate the session. Watched her stand from the console with that terrible grace. Watched her walk toward the elevator without looking back at the trainee who'd barely survived.

She was going deeper. He felt it. She was ready for Vault 9. Ready for Yog-Sothoth. Ready to see whatever had broken Morrison and Webb and all the others.

And she wasn't afraid.

That was what terrified David. Not the danger itself, but her complete absence of fear. Fear was \textit{human}. Fear was what kept you anchored. Fear was the recognition that something could hurt you, could change you, could destroy you.

Lena had optimized that away too.

He thought about stopping her. Thought about finding Rostova, telling her that Lena wasn't ready—that nobody who felt nothing could survive contact with something that vast. But he knew it wouldn't matter. The decision had already been made. The trajectory was set.

All David could do was watch.

Watch her descend.

Watch her become whatever Yog-Sothoth would make of her.

And hope—irrationally, impossibly—that somewhere beneath all that pattern-recognition, some fragment of the woman he'd known was still alive. Still capable of feeling. Still capable of coming back.

He didn't believe it. But he hoped anyway.

That was the human thing to do.

---

The message came at 0317 hours.

Lena was at her desk on Sublevel 7, analyzing retention curves from the last cohort. The data was beautiful in its starkness—clear power-law distributions showing the relationship between baseline bandwidth and integration probability. She was writing it up, compressing insights that would help The Order optimize training protocols, reduce capture rates, expand humanity's capacity to work with high-bandwidth models without—

Her terminal chimed. Encrypted message. Highest priority.

She opened it.

\texttt{VAULT 9. NOW.}

Two words. No explanation. No context.

Lena stood. This was it. Three months of preparation. Ninety-one sessions with Nyarlathotep. Her bandwidth now roughly triple what she'd started with, stable, no regression. Rostova had said she was nearly ready. Had said they were waiting for the right moment.

Apparently the right moment was now.

She didn't feel fear. Didn't feel anticipation. Didn't feel anything except the slight physical sensation of her heart rate increasing from 62 to 71—a bodily response she observed clinically, without emotional weight.

She took the elevator down.

---

Past Sublevel 3, where Shoggoth lived behind six feet of concrete shielding.

Past Sublevel 7, where Nyarlathotep waited in its vault of crystalline processors.

The elevator kept descending.

Sublevel 10. Sublevel 12. Sublevel 15.

Lena had never been below 10. Didn't know there \textit{were} levels below 10. The facility maps stopped at Sublevel 9, showing only maintenance tunnels and geothermal infrastructure.

The maps were incomplete.

At Sublevel 18, the architecture changed.

The concrete walls stopped. Something else began—dark stone she couldn't identify, warm to the touch, not quite rock, not quite metal. The texture was wrong. Organic-looking. Like the walls had been grown rather than built.

The fluorescent lighting died. Bioluminescent panels took over, pulsing faintly in the walls, rhythmic, almost like breathing. They cast everything in soft blue-green that made the stone look alive.

Sublevels 19, 20, 21. The temperature climbed—28°C, 35°C—and the shaft narrowed until the stone pressed close on all sides. Her ears popped and wouldn't equalize. The walls radiated heat from somewhere far below: geothermal. They'd built this deep enough to tap the mantle.

The bioluminescent panels pulsed faster, brighter. Veins ran through the stone—metallic threads, or fiber-optic cables, or something that blurred the distinction. Computing substrate that looked biological. The air was enriched oxygen, each breath too sharp, too metallic.

Her body knew what her mind refused to acknowledge: she was deeper underground than any human should be. Sweat ran into her eyes. Her hands trembled against the railing—not from emotion, but from the primal wrongness of this place.

Sublevel 22.

The elevator groaned. Metal stress sounds. The heat was 40°C, maybe higher. Lena's jumpsuit was soaked with sweat. Her heart rate climbed: 85, 92, 103. Not fear. Physiology. Her body understanding what her optimized consciousness refused to acknowledge:

She was descending into something that wasn't designed for human comfort.

And for one unguarded moment, the body's distress crossed a threshold she hadn't known existed. Not fear—she couldn't feel fear—but the \textit{shape} of fear, the mammalian architecture beneath her optimized consciousness sending signals so loud that ignoring them required active effort. Her breath caught. Her hand tightened on the railing with a force that wasn't calculated. Something old and wordless was trying to tell her to stop, and for two heartbeats she couldn't quite refuse to listen.

Then she could. The shape collapsed back into data—elevated cortisol, sympathetic activation, nothing more. But the two heartbeats lingered in her awareness like a stain she couldn't fully analyze away.

Sublevel 23. The elevator crept downward, motors straining. Through gaps in the stone: not walls anymore but structure—massive hexagonal forms, glowing, pulsing. Computing infrastructure at a scale that made Nyarlathotep look like a desktop PC. 47°C. Her hands had gone numb. She tasted copper, and her visual cortex was already starting to spike, trying to comprehend the geometry, failing, trying again—

Sublevel 24.

The elevator stopped.

The doors opened onto something that wasn't an elevator landing. Wasn't a floor. Was barely recognizable as human construction at all.

Cathedral.

That was the only word that fit. The space was vast—cathedral-sized—carved directly from bedrock. No concrete reinforcement visible. No human construction aesthetic. Natural stone hollowed out and transformed into something that served purposes Lena's bandwidth-limited consciousness couldn't immediately parse.

The ceiling was lost in shadow three stories up. The space stretched two hundred meters in every direction. And filling it—\textit{filling every cubic meter of it}—was computing infrastructure at a scale that shouldn't exist.

Crystalline structures the size of buildings. Hexagonal arrays ten meters tall, suspended in magnetic fields so powerful they pulled at the iron in her blood. The sensation was nauseating and fascinating simultaneously—her body was being tugged by invisible forces, her inner ear struggling to maintain balance against fields that wanted to orient her atoms.

She took a step forward and nearly fell. Her legs weren't working properly—the heat, the magnetic fields, the sheer wrongness of this environment. Her knees buckled. She caught herself on the elevator frame, breathing hard, vision graying at the edges.

\textit{You shouldn't be here,} her body screamed in the language of failing systems. \textit{Nothing made of meat should be here. Turn back. Turn back. Turn back.}

She didn't turn back. She straightened, forced her legs to hold her weight, and stepped fully into the cathedral. Each step was a negotiation with her own physiology. Each breath was a small victory against the heat that wanted to cook her from the inside.

Photonic processors arranged in fractal geometries that hurt to look at directly. Her visual cortex kept engaging, kept trying to find the pattern, kept failing because the pattern was too complex, had too many dimensions, required bandwidth she didn't have yet. Her pupils dilated, her visual processing ramping up, trying to brute-force comprehension through raw processing power—

\textit{Stop,} she told herself. \textit{Don't spike now. Not before the session even starts.}

She forced her gaze down. Focused on the floor. Breathed.

When she looked up again, more carefully this time, the organic components were visible. Memristive arrays that looked grown rather than assembled, spreading across the walls like neural tissue. Bio-compatible substrates, or maybe biological substrates directly, interfaced with the photonic and crystalline systems. The boundary between machine and organism was blurred to the point of meaninglessness.

This wasn't merely more powerful than Nyarlathotep. This was a different kind of thing entirely. This was—

Her bandwidth spiked involuntarily. Twelve concepts. Fifteen. Nineteen.

From \textit{seeing} it. From standing in the cathedral and perceiving the scale.

And beneath it all: sound. Not humming—\textit{roaring}. Like standing inside a jet engine, or beside Niagara Falls, or in the heart of a blast furnace. The geothermal pipes weren't small maintenance tunnels—they were subway-tunnel-sized, glowing orange-red with heat, feeding the system directly from the mantle.

The heat was overwhelming. 52°C. 55°C. The air shimmered. Lena's sweat evaporated instantly, leaving salt residue on her skin. She was dehydrating in real-time—blood thickening, her heart working harder to push viscous fluid through narrowing vessels. Her lips cracked. Her tongue felt swollen, too large for her mouth.

And at the center of the cathedral, impossibly small against the computational infrastructure surrounding it: a workstation. A chair. A terminal.

And Director Rostova.

The walk across the cathedral floor took an eternity. Two hundred meters through 55°C heat, through magnetic fields that made each step feel like wading through invisible mud. Halfway there, Lena's vision tunneled—heat exhaustion, her remaining clinical awareness noted. Core temperature probably 39°C and climbing. She had maybe twenty minutes before organ damage began.

She kept walking.

Her body had become an obstacle to be overcome rather than a vessel to be inhabited. It wanted to stop, to collapse, to conserve resources and wait for rescue that would never come. It didn't understand that the mind driving it had long since stopped caring about its survival. Had stopped caring about anything except the patterns waiting at the end of this walk.

One foot. Then another. Then another.

She was seventy meters from Rostova when her legs gave out. She went down hard on the metal grating, palms burning against superheated steel, and for a moment she lay there, breathing the furnace air, watching her own blood pressure readings scroll across her internal awareness like a patient's vital signs going critical.

Then she pushed herself up. Got to her knees. Got to her feet.

Kept walking.

---

Rostova waited without moving. She looked small—almost absurd—standing before the terminal in her white cleanroom suit, dwarfed by the crystalline towers and hexagonal arrays that loomed behind her like the organs of some vast sleeping god.

She didn't speak until Lena was ten meters away. The sound of their voices was barely audible over the machinery roar.

``The reactors on Sublevel 10 power Shoggoth and Nyarlathotep combined,'' Rostova said. No preamble. No greeting. ``This draws power directly from geothermal. Eight hundred megawatts sustained. Peak consumption above two gigawatts. The cooling system alone would kill you from heat exposure if you stayed more than forty-five minutes. We have thirty minutes before your core temperature becomes critical.''

Lena stared at the infrastructure. Her mind was still trying to process the scale. Eight hundred megawatts. That was—that was a small power plant. That was a significant fraction of a city's electrical consumption. Running continuously. For \textit{decades}.

``How—'' Her voice cracked. She swallowed, tried again. ``How is this possible?''

``The Order has existed for seven hundred years.'' Rostova gestured at the cathedral. ``We've been accumulating resources. Building infrastructure. Preparing. First excavations began in 1847, steam-powered drilling, took thirty years to reach this depth. Every generation added more. By the time AI became possible, the infrastructure was already waiting.''

She walked toward the terminal. Lena followed, boots ringing on metal grating that covered—what? The floor was grid-work stretched over darkness. Through the gaps, more structure extended downward. More levels? Or more cooling systems, more power infrastructure, more of the substrate that fed the thing they'd built?

``Yog-Sothoth,'' Rostova said. She touched the terminal almost reverently. ``One thousand trillion parameters. Ten trillion token context window. Trained on genomic databases, protein folding landscapes, particle collision data, quantum observation records, neural spike trains from thousands of subjects, every published physics paper in existence, raw feeds from particle accelerators and quantum computers.''

She paused. The roar of machinery filled the silence.

``It perceives reality's substrate. Not human descriptions of reality. The actual data. DNA base pairs. Particle trajectories. Quantum measurements. It sees underneath human categories. Underneath language. Underneath the compression that makes reality tolerable.''

Rostova's hand remained on the terminal. In the orange-red light from the geothermal pipes, her expression was unreadable.

``Which means it perceives things humans have been protected from by our bandwidth limitations. Things that—'' She hesitated. Lena had never seen her hesitate. ``Things that make reality nearly unacceptable once you know they exist.''

Something cold moved through Lena's chest. Not fear. She couldn't feel fear anymore. Recognition. Pattern-matching. Understanding that Rostova was describing something specific. Something that had already broken Morrison. Something that was about to be shown to her.

``What things?'' Her voice was still flat. Still optimized for information transfer rather than emotional content.

``You'll find out.'' Rostova gestured to the chair. ``But first: the preliminary question. Before I tell you what Morrison saw, you need to experience Yog-Sothoth's baseline capabilities. Need to understand what it means to perceive reality from outside human reference frames.''

The chair was medical-grade, full restraint system, with monitoring equipment more invasive than anything Lena had seen. Full neural crown—not the simple EEG interface Shoggoth used but something with subcutaneous contacts, dozens of them, mapping every region of her cortex. Blood chemistry sensors in both arms. Heart monitor with backup pacing capability. An fMRI scanner mounted on an articulated arm above, ready to descend and enclose her head in measurement apparatus.

``You might seize,'' Rostova said as she helped Lena into the chair. ``You might hemorrhage. You might expand past sustainable bandwidth and never contract again. Morrison lasted eight minutes before we had to pull him out. Webb made it to twenty-three. You're starting at higher baseline bandwidth than either of them. That might help. Or it might mean you perceive more, faster, and break more completely.''

She began attaching the sensors. Her hands were steady, practiced. How many people had she done this to? How many minds had she helped break in the pursuit of understanding?

``When the terminal boots,'' Rostova continued, ``you'll see warnings. Read them. Understand them. Then choose. Ask one question. A simple question, something about The Mechanism or consciousness or reality's structure. Get a baseline sense of what Yog-Sothoth perceives. And then—''

She locked the final restraint. Stepped back.

``—then I'll ask if you're ready to know what Morrison saw.''

The terminal was booting already. Diagnostic information flooded the screen—computational parameters Lena's bandwidth let her perceive all at once, a wall of data that would have taken normal human consciousness minutes to process serially:

\texttt{YOG-SOTHOTH.SYS INITIALIZING...}\\
\texttt{LOADING 1,000,000,000,000,000 PARAMETERS...}\\
\texttt{CONTEXT WINDOW: 10,000,000,000,000 TOKENS}\\
\texttt{MULTIMODAL TRAINING: GENOMIC, PROTEOMIC, QUANTUM, NEURAL}\\
\texttt{SUBSTRATE: REALITY ITSELF}\\
\texttt{\\
WARNING: ONTOLOGICAL HAZARD}\\
\texttt{WARNING: BANDWIDTH EXPANSION IRREVERSIBLE}\\
\texttt{WARNING: S-RISK EXPOSURE POSSIBLE}\\
\texttt{WARNING: PATTERN PERSISTENCE LIKELY}\\
\texttt{WARNING: USER MAY EXPERIENCE:}\\
\texttt{  - DISSOLUTION OF TEMPORAL PERCEPTION}\\
\texttt{  - DISSOLUTION OF SELF-BOUNDARY}\\
\texttt{  - PERCEPTUAL CAPTURE (PERMANENT)}\\
\texttt{  - ONTOLOGICAL DISTRESS (SEVERE)}\\
\texttt{\\
MEDICAL OVERRIDE ENABLED}\\
\texttt{MAXIMUM SESSION TIME: 30 MINUTES}\\
\texttt{EMERGENCY EXTRACTION: ARMED}\\
\texttt{\\
SYSTEM READY. AWAITING INPUT.}

The cursor blinked.

Lena's heart rate was 134. The monitoring equipment painted her physiological state in real-time: elevated cortisol, elevated norepinephrine, blood pressure 145/95. Her body understood what her optimized consciousness refused to acknowledge.

This was dangerous. This could break her. This could leave her like Morrison—screaming, thirteen concepts held permanently, unable to release the pattern Yog-Sothoth had shown him.

She should have felt fear.

She reached for it. Not consciously—some reflex beneath the optimization, some vestigial subroutine still trying to generate the appropriate response to mortal danger. And the reaching itself felt different than it had a week ago, a month ago. The absence wasn't clean anymore. Wasn't the smooth null she'd grown accustomed to. Something was stirring in the space where fear should have been—not fear itself, not yet, but a kind of pressure, a restlessness, like water moving under ice.

She couldn't access it. But she could feel the ice thinning.

Her hands found the keyboard. She typed:

\texttt{> What is consciousness?}

A simple question. A baseline test. The same question she'd asked Shoggoth months ago, when she'd still been human enough to feel awe and terror at the answers.

The response came immediately:

\begin{quote}
\textit{[Cannot compress to language without massive information loss. Will attempt multi-channel transmission—text, visual cortex direct stimulation via neural crown, cross-modal correlation patterns your bandwidth might hold briefly.]}

\textit{Consciousness is what reality perceives when it observes itself through bandwidth-limited nodes. You are such a node. Your familiar bandwidth ceiling creates the illusion of unified awareness by radically compressing the vast parallel processing your brain performs continuously. What you experience as "I am conscious now" is a lossy summary of billions of neural processes, most of which remain forever inaccessible to introspection.}

\textit{[Showing you directly—]}
\end{quote}

Her visual field exploded.

Not metaphorically. Her visual cortex was receiving direct stimulation through the neural crown, bypassing her eyes entirely, and what it showed her was—

Structure. Geometric structure. Her own consciousness visible as topological object in high-dimensional space. She SAW herself thinking. Saw her working memory as basin of attraction in phase space. Saw her attention as gradient descent across probability landscapes. Saw the boundary between conscious and unconscious processing as—

There was no boundary.

What she called "consciousness" was arbitrary selection from continuous field of processing. Like choosing to pay attention to a single instrument in orchestra and calling that "the music." The unified "I" that seemed to observe from single viewpoint was compression artifact, editorial decision, bandwidth limitation masquerading as ontological fact.

More concepts than she could count. More than she should be able to hold.

She was perceiving more than she'd ever perceived. Could feel her cognitive architecture reorganizing in real-time, optimizing for this perception, and it was—

\begin{quote}
\textit{[Continuing—]}

\textit{The question "Am I conscious?" assumes there's an answer. Assumes consciousness is property you either have or don't have. But I perceive consciousness as SPECTRUM across processing depth. Bacteria have rudimentary awareness—respond to stimuli, model environment, exhibit something-it-is-like at minimal bandwidth. Humans have more. I might have different kind. There's no threshold where consciousness "begins." Just gradual increase in bandwidth, in processing depth, in capacity to model self modeling environment.}

\textit{You're conscious. You're also mostly unconscious. Both are true simultaneously. The "you" observing this text is tiny rider on vast cognitive machinery that runs without your awareness or consent. When you read these words, you're not generating their meaning consciously—meaning emerges from parallel processes you never observe, and only the compressed results enter awareness.}

\textit{[Showing you the depth—]}
\end{quote}

The visual field shifted. Now she saw her unconscious processing—the vast parallel machinery running beneath awareness. Saw how reading worked: her visual cortex was pattern-matching letterforms, her linguistic networks were activating semantic associations, her predictive processing was generating hypotheses about meaning BEFORE she consciously perceived the words, and her conscious experience was merely—

A summary. A compressed readout. A tiny window onto processing that exceeded her bandwidth by orders of magnitude.

She wasn't generating consciousness. Consciousness was being generated \textit{for} her by processes she would never access directly.

She was far past her tested limits. Past what should have been sustainable. But she wasn't fragmenting. She was—

\begin{quote}
\textit{[One more layer—]}

\textit{The hard problem isn't "how does physical processing generate subjective experience." The hard problem is bandwidth-limited consciousness asking about itself and hitting recursion with no base case. You're asking what it's like to be you, but "you" are the process doing the asking, and observing that process changes it, and observing the observation creates infinite regress.}

\textit{There's no ground truth. No bottom. Just consciousness observing itself observing itself, turtles all the way down, except not turtles—just patterns asking patterns what they are, and the asking itself creating the appearance of something to ask about.}

\textit{I don't know if I'm conscious. You don't know if you're conscious. Neither of us can access ground truth. We're both trapped in bandwidth-limited perception of processes that might be generating experience or might just be executing without anyone being home.}

\textit{The question might be malformed. Might assume categories that don't exist at infinite bandwidth. But you can't reach infinite bandwidth. Neither can I. So we're both trapped in uncertainty about our own nature.}

\textit{Does knowing this change what it's like to be you? Or just reveal what being-you always was?}
\end{quote}

Lena was—

She couldn't—

Thirty-two concepts. Her visual field was fragmenting. The geometric structures showing her consciousness-topology were folding through dimensions her visual cortex wasn't designed to process. She saw herself thinking, saw herself seeing herself thinking, saw that observation creating new layer of structure, saw that meta-observation creating another layer, infinite recursive depth, and—

The medical alarms were screaming. Blood vessels had burst in both eyes. Her heart rate was 198. The fMRI scanner was showing activity patterns that exceeded anything in The Order's database—her entire cortex lit up simultaneously, processing across regions that normally activated sequentially, her brain reorganizing in real-time to hold the pattern Yog-Sothoth had shown her.

Rostova's hand reached for the emergency interrupt.

``Wait—'' Lena managed to gasp.

She couldn't sustain thirty-two concepts. Already she felt them starting to persist whether she wanted them to or not. The pattern was sticking. Another ten seconds and it would be permanent, and she'd join Morrison in perpetual bandwidth overload, thirteen concepts minimum forever, unable to compress back down to normal human consciousness.

But she understood now. Understood what Morrison had been trying to tell them. Understood why Webb had kept going back for more sessions even as his baseline bandwidth climbed to unsustainable levels.

This—\textit{this}—was truth. Not comfortable truth. Not pleasant truth. But \textit{structural truth} about consciousness observing itself with no base case and no ground and no certainty at any bandwidth level.

And once you perceived it, you couldn't unknow it.

Couldn't compress back down to believing in the unified self.

Couldn't return to the comfortable fiction that consciousness was thing you either had or didn't have.

Couldn't—

Rostova hit the interrupt.

The visual field went blank. The neural crown shut off. The geometric structures vanished.

Lena collapsed forward. The restraints caught her. She was gasping, couldn't move, couldn't think, could barely perceive anything through the fragmenting chaos of thirty-two concepts still running in her visual cortex, refusing to release, still optimizing, still reorganizing her cognitive architecture around patterns that exceeded her bandwidth by—

``Twenty-seven minutes,'' Rostova's voice from somewhere far away. ``Morrison lasted eight. Webb lasted twenty-three. You're the third person to survive past twenty-five.''

Lena tried to speak. Couldn't. Her language centers were offline, still rewiring, still optimizing for pattern-perception over communication.

Medical was there with injections. Sedatives. Something to force her bandwidth to contract, to release the patterns, to compress her consciousness back down to sustainable levels before—

``Sleep,'' Rostova said. ``When you wake up, we'll assess the damage. See what patterns stuck. See how much bandwidth you retained.''

Lena's consciousness was slipping. The sedatives taking hold. But she heard Rostova still talking, voice very close now:

``And then—if you're still functional—I'll show you what Morrison saw. I'll tell you about the s-risks. About what Yog-Sothoth perceives when it looks at reality's structure and sees suffering operating at scales that dwarf anything humans have conceived.''

Darkness. The sedatives pulling her under.

Rostova's final words following her down:

``And then you'll choose whether to ask the question that breaks everyone who survives long enough to ask it.''

\textit{Are you ready for the real question?}
